Automatic Detection of Bad Programming Habits in Scratch: A Preliminary Study

### 2014-04-23 paperFIE_en.tex
Add results table. Add comment regarding length of programs to be considered equal.

However, there are reports that indicate that students learning to program in this environment show certain habits that are contrary to the basic programming recommendations. In our work as instructors with high school students, we have detected these and other bad practices, such as the repetition of code and object naming, on a regular basis. This paper focuses on verifying whether these issues can be generally be found in the Scratch community, by analyzing a large number of projects available on the Scractch community website.

@@

  \item \textit{duplicate.DuplicateScripts}\footnote{\url{https://github.com/jemole/hairball/blob/master/hairball/plugins/duplicate.py}} analyzes a Scratch project to try to locate repeated complete programs within a project. In order to do so, the plug-in runs each project and gets the \textit{tokens} of the different blocks of the programs, so that two blocks that only differ in the values they receive are considered to be equal. To illustrate the operation of this plug-in, Fig.~\ref{fig:CodeRepetition1} shows two scripts containing the same blocks with the only difference of the values received by parameter.

@@

The results obtained after analyzing the 100 projects confirm that observations made with our students can also be generally found in the community projects.

@@

>>>>

However, there are reports that indicate that students learning to program in this environment show certain habits that are contrary to the basic programming recommendations. In our work as instructors with high school students, we have detected these and other bad practices, such as the repetition of code and object naming, on a regular basis. This paper focuses on verifying whether these issues can generally be found in the Scratch community, by analyzing a large number of projects available on the Scractch community website.

@@

  \item \textit{duplicate.DuplicateScripts}\footnote{\url{https://github.com/jemole/hairball/blob/master/hairball/plugins/duplicate.py}} analyzes a Scratch project to try to locate repeated complete programs within a project. In order to do so, the plug-in runs each script of the project and gets the \textit{tokens} of the different blocks of the programs, so that two blocks that only differ in the values they receive are considered to be equal. A minimum length of five blocks is applied for programs to be considered duplicate rather than coincidentally similar.

  To illustrate the operation of this plug-in, Fig.~\ref{fig:CodeRepetition1} shows two scripts containing the same blocks with the only difference of the values received by parameter.

@@

The results obtained after analyzing the 100 projects confirm that observations made with our students can also be generally found in the community projects, as shown in table \ref{table:results}.

@@

\begin{table}
\begin{center}
  \begin{tabular}{ | p{2.5cm} | p{1cm} | p{1.2cm} | p{1cm} |}
   \hline
     & Default names & Duplicated scripts & Defined blocks \\ \hline\hline
    Projects incurring in & 79 & 62 & 17 \\ \hline
    Mean & 5.94 & 7.23 & 1.11 \\ \hline
    Median & 3 & 2 & 0 \\ \hline
    Maximum & 67 & 71 & 25 \\
    \hline    
  \end{tabular}
\end{center}
\caption{Table of results}
\label{table:results}
\end{table}
